---
title: "Exact Simulation Results"
subtitle: "Shiny App developed by Dani&euml;l Lakens and Aaron Caldwell"
output: pdf_document
date: '`r format(Sys.time(), "%B %d, %Y")`'
params:
  tablePC: NA,
  tableMain: NA,
  means_plot: NA,
  n: NA,
  model: NA,
  design: NA,
  cor_mat: NA,
  sigmatrix: NA
  alpha_level: NA
  power_curve: NA
  power_curve_df: NA
---




Below are the results from a power analysis using an exact simulation of a factorial design wherein the data are sampled from an empircal distribution. If you encounter any problems please visit our GitHub page (https://github.com/Lakens/ANOVApower) to the raise the issue.

**Study Design**

`r paste(params$design)`

**Model Formula** 

`r paste(params$model)`

The sample size was **`r params$n`** *per cell*.

**Correlation Matrix**
```{r echo = FALSE}
knitr::kable(params$cor_mat)
```

**Variance-Covariance Matrix**
```{r echo = FALSE}
knitr::kable(params$sigmatrix)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Plot of the individual data points with mean and standard deviation.

```{r echo = FALSE}
params$means_plot
```


```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Power for this specific design, with an alpha of `r params$alpha_level`, is included below.

**ANOVA Power (%) and Effect Sizes (Partial Eta Squared)**

```{r echo = FALSE}
knitr::kable(params$tableMain)

```

**Multiple Comparisons Power (%) and Effect Sizes (Cohen's d~z~)**

```{r echo = FALSE}
knitr::kable(params$tablePC)

```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

**Plot of Power Across Many Sample Sizes**

```{r echo = FALSE}
params$power_curve
```



```{r, results='asis', echo=FALSE}
cat("\\newpage")
```


**Table of Power Across Many Sample Sizes**

```{r echo = FALSE}
knitr::kable(params$power_curve_df)
```
