---
title: "Exact Simulation Results"
subtitle: "Shiny App developed by Dani&euml;l Lakens and Aaron Caldwell"
output: pdf_document

date: '`r format(Sys.time(), "%B %d, %Y")`'
params:
  tablePC: NA,
  tableMain: NA,
  means_plot: NA,
  n: NA,
  model: NA,
  design: NA,
  cor_mat: NA,
  sigmatrix: NA
  alpha_level: NA
  tableEMM: NA
  input_emm: NA
---


```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
options(knitr.table.format = 'markdown')
```


Below are the results from a power analysis using an exact simulation of a factorial design wherein the data are sampled from an empircal distribution. If you encounter any problems please visit our GitHub page (https://github.com/arcaldwell49/Superpower) to the raise the issue.

**Study Design**

`r paste(params$design)`

**Model Formula** 

`r paste(params$model)`

The sample size was **`r params$n`** *per cell*.

**Correlation Matrix**
```{r echo = FALSE}
knitr::kable(params$cor_mat)
```

**Variance-Covariance Matrix**
```{r echo = FALSE}
knitr::kable(params$sigmatrix)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Plot of the design with mean and standard deviation.

```{r echo = FALSE}
params$means_plot
```


```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Power for this specific design, with an alpha of `r params$alpha_level`, is included below.

**ANOVA Power (%) and Effect Sizes (Partial Eta Squared)**

```{r echo = FALSE}
knitr::kable(params$tableMain)

```

**Estimated Marginal Means Comparisons Power (%) and Effect Sizes (Cohen's *f*)**

If estimated marginal means were selected, the results are included below. 

```{r echo = FALSE}
if(params$input_emm == "yes"){
knitr::kable(params$tableEMM)
}
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

**Multiple Comparisons,*t*-test, Power (%) and Effect Sizes (Cohen's d~z~)**

Below are the power analysis results if all pairwise comparisons are completed via *t*-tests

```{r echo = FALSE}
knitr::kable(params$tablePC)
```




