---
title: "Power Calculations for ANCOVA"
author: "Aaron R. Caldwell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: refs.json
link-citations: true
output:
  rmarkdown::html_vignette:
    df_print: "kable"
    toc: true
vignette: >
  %\VignetteIndexEntry{Power Calculations for ANCOVA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  # collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library("Superpower")
```

# Introduction

Recently, @shieh_ancova demonstrated that most software uses a slightly flawed approach to estimating power for ANCOVAs, and focused on one-way ANOVA designs to compare his method to the method first mentioned by @cohen_book.

# One-way ANOVA

As @shieh_ancova eloquently points out in their simulations, the method of overestimates power on 

# Extending to factorial ANOVAs

# Contrasts

```{r}
ANCOVA_contrast(cmat = c(-1,1),
                n = 15,
                mu = c(0,1),
                sd = 1,
                r2 = .2,
                n_cov = 1)
```


# Method

# References