---
title: "Monte Carlo Simulation Results"
subtitle: "Shiny App developed by Dani&euml;l Lakens and Aaron Caldwell"
output: pdf_document
date: '`r format(Sys.time(), "%B %d, %Y")`'
params:
  tablePC: NA,
  tableMain: NA,
  tableEMM: NA,
  means_plot: NA,
  n: NA,
  model: NA,
  design: NA,
  cor_mat: NA,
  sigmatrix: NA
  alpha_level: NA
  nsims: NA
  p_adjust: NA
  correction: NA
  manova: NA
  emm_p_adjust: NA
  emm: NA
  emm_model: NA


---

```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
options(knitr.table.format = 'markdown')
```


Below are the results from a power analysis using a Monte Carlo simulation of a factorial design wherein the data are sampled from a multivariate normal distribution. If you encounter any problems please visit our GitHub page (https://github.com/arcaldwell49/Superpower) to the raise the issue.

**Study Design**

`r paste(params$design)`

**Model Formula** 

`r paste(params$model)`

The sample size was **`r params$n`** *per cell* with a total number of observations of **`r params$n*length(diag(as.matrix(params$sigmatrix)))`**.

**Correlation Matrix**
```{r echo = FALSE}
knitr::kable(params$cor_mat)
```

**Variance-Covariance Matrix**
```{r echo = FALSE}
knitr::kable(params$sigmatrix)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

Plot of the individual data points with mean and standard deviation.

```{r echo = FALSE}
params$means_plot
```


```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

The results of the Monte Carlo simulation are included below. This simulation assumed an alpha of **`r params$alpha_level`**, a sphericity correction of **`r params$correction`**, a **`r params$p_adjust`** adjustment for multiple comparisons. A total of **`r params$nsims`** iterations were included in this simulation.

**ANOVA Power (%) and Effect Sizes (Partial Eta Squared)**

```{r echo = FALSE}
knitr::kable(params$tableMain)

```

**Estimated Marginal Means Power (%) and Effect Sizes (Cohen's *f*)**

*Only included if the estimated marginal means were selected.*

`r if(params$emm == "yes"){paste("For the estimated marginal means comparisons, the", params$emm_model," model was utilzed with the following correction for multiple comparisons:", params$emm_p_adjust)}`


```{r echo = FALSE}
if(params$emm == "yes"){
knitr::kable(params$tableEMM)
}
```


**Power for MANOVA Results**

*Only included if there is a within-subject factor.*

```{r echo = FALSE}
if(!is.null(params$manova)){
knitr::kable(params$manova)
}
```

**Multiple Comparisons, *t*-tests, Power (%) and Effect Sizes (Cohen's d~z~)**

```{r echo = FALSE}
knitr::kable(params$tablePC)
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

